### 2023-02-03 article3.qmd
update

depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, NP_Size_512) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
  mutate(month = month.abb[month(juld)]) |> ungroup() #|> filter(month != 'May') # I filtered May because I had some troubles.. I need to fill some cases unseen by the float (eg. when they start profiling until 1000 m then go to 2000 -> ça fout un peu la merde..)

tmp <- part_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_part = log10(mean(NP_Size_512))) |> 
  mutate(cut_part = cut(mean_part, breaks = c(seq(0,100,20),+Inf))) |> ungroup()

@@

ggplot(dd) + 
  geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Flux range', x = 'Month', y = 'Depth bin') +
  facet_wrap(~wmo) + scale_fill_viridis(option = 'A')

@@

>>>>

depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, NP_Size_256) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
  mutate(month = month.abb[month(juld)]) |> ungroup()

tmp <- part_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_part = log10(mean(NP_Size_256))) |> 
  mutate(cut_part = cut(mean_part, breaks = c(seq(0,100,20),+Inf))) |> ungroup()

@@

ggplot(dd) + 
  geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Log10(abundance) #/L', x = 'Month', y = 'Depth bin', title = 'NP_Size_256') +
  facet_wrap(~wmo) + scale_fill_viridis(option = 'A')

#### same plots mais par mois

Il y a une erreur lors de la standardisation -> WHY doesn't it want to group it by month???

```{r, monthly heatmap}

depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, all_of(lpm_class)) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |> mutate(month = month.abb[month(juld)]) |> ungroup() |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration')

tmp <- part_distrib |> group_by(depth_bin, size, month) |> summarize(mean_part = mean(concentration)) |> mutate(rel_mean_conc = (mean_part-min(mean_part))/(max(mean_part) - min(mean_part))) |> ungroup()

plot_heat_month <- function(m){
  # transform data for heatmap
  #m <- 'May'
  tmp2 <- tmp |> filter(month == m) |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric()) |> select(-c(month, mean_part))
  y <- unique(tmp2$depth_bin)
  x <- unique(tmp2$size)
  test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = rel_mean_conc) 
  #test <- tmp2 |> pivot_wider(names_from = depth_bin, values_from = mean_part) 
  data <- expand.grid(X=x, Y=y)
  data$Z <- as.vector(as.matrix(test |> select(-size)))
  data$month <- m
  return(data)
}

m_data <- map_dfr(c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'), plot_heat_month)

m_data <- m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)),
                   month = factor(month, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')))

ggplot(m_data |> mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T)))) + 
  geom_tile(aes(as.factor(X), Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + scale_fill_viridis(option = 'A') +
  facet_wrap(~month)

@@

```
### Modèle basé sur les pentes

```{r, model based on slope}

# compute biovolumes for mid size bin (ESD)
# DSE from 102 microns to 2.05 mm (this class to goes to 2.58 mm) (values obtained from Guidi's Matlab code converted on python)
DSE <- c(0.11525597, 0.14521343, 0.18295745, 0.23051195, 0.29042685,
              0.36591491, 0.46102389, 0.58085371, 0.73182981, 0.92204779,
              1.16170742, 1.46365963, 1.84409558, 2.32341484)

# DSE obtained with Rainer's code
DSE_Rainer <- c(0.11645117591608793,0.14636003747903648,0.18439151433041295,
                0.23251999827981656,0.2933257316240555,0.36916500651202966,0.4650399965596331,
                0.5860454935035602,0.738553273391974,0.9280422447589767,1.1705684285914486,
                1.4795320531800742,1.8636618038457486,2.3449457391257185)

# compute biovolume
get_biovolume <- function(DSE){
  r <- DSE/2
  biovolume <- (4*pi/3)*r**3
  return(biovolume)
}

# get biovolume for each DSE (equivalent to say for each size bin)
biovolumes <- map_dbl(DSE, get_biovolume)  # unit = m^3

# spectra at parking depth
uvp_spectra <- all_floats |> filter(PhaseName == 'PAR') |> mutate(month = month(juld), week = week(juld), DOY = yday(juld), short_date = ymd(juld)) |> select(juld, month, week, DOY, short_date,  cycle, wmo, park_depth, all_of(lpm_class)) |> 
  drop_na(NP_Size_102) |> mutate(juld = as_date(juld), WMO = factor(wmo))

# convert abundance to biovolume
spectra_biovolume <- tibble(as.data.frame(t(t(uvp_spectra[,all_of(lpm_class)])*biovolumes))) 
colnames(spectra_biovolume) <- colnames(spectra_biovolume)|> str_replace('Size', 'Biovolume')

uvp_spectra <- cbind(uvp_spectra, spectra_biovolume) |> mutate()

cflux <- cflux |> mutate(small_slope = (small_flux/633)**(1/0.77), big_slope = (big_flux/633)**(1/0.77))

## try model for small slopes
cflux2 <- cflux |> mutate(flux = big_slope)

data_model <- merge(cflux2, uvp_spectra) |> mutate(depth = if_else(park_depth == '1000 m', 1000, if_else(park_depth == '500 m', 500, 200))) |> select(-park_depth)

# training data
tmp <- data_model |> select(flux, all_of(colnames(spectra_biovolume)))

set.seed(101)
boosting_rounds <- 3000
# model
model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
cv.folds=5,
n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
bag.fraction=0.5,
n.cores=1
)
